Prospects of detecting gamma-ray emission from galaxy clusters: cosmic rays and dark matter annihilations

### 2011-04-09 paper.tex
updated

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
 \includegraphics[width=0.49\columnwidth]{figures/flux.BMcompI.v9.0.1deg.1.6T.SubMass.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.BMcompJ.v9.0.1deg.1.6T.SubMass.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.BMcompK.v9.0.1deg.1.6T.SubMass.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.BMcompJs.v9.0.1deg.1.6T.SubMass.IR2.noMW.woGal.eps}
\caption{\it Comparing the differential flux from different models: we
  show the continuum emission from DM benchmark (BM) models (blue
  dashed), electrons and positrons from DM BM models that inverse
  Compton upscattered both CMB photons (orange dash-dotted) as well as
  dust and star photons (green dotted), and CRp induced gamma-ray
  emission (red solid). Each panel is associated with an individual DM
  BM model; upper left $\Ip$, lower left $\Kp$, upper right $\Jp$, and
  lower right $\Js$. The emission is calculated for the Fornax cluster
  using a point spread function of $0.1\deg$ and a boost from
  substructures of 570. We find bright prospects for detecting the
  continuum emission that is dominating the total emission in the GeV
  energy range. Also note that, above 100~MeV the total inverse
  Compton emission is at least a factor 10 smaller than the emission
  expected from both the continuum emission and the emission coming
  from CRps.}
 \label{fig:diff_BM}
\end{minipage}
\end{figure*}

@@

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
\includegraphics[width=0.49\columnwidth]{figures/flux.IRcomp.v9.0.1deg.1.6T.SubMass.elmu.SF300.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.IRcomp.BMv9.0.1deg.SubMass.noMW.woGal.eps}
\caption{\it Comparing the flux from different inverse Compton
  upscattered radiation fields. We show the differential inverse
  Compton emission induced by leptophilic DM in the left panel and by
  the $\Kp$ benchmark model in the right panel. The contribution from
  each individual radiation field from top line to bottom line; CMB
  (green dashed-double-dotted), dust (blue dashed), stars (purple
  dashed-dotted). The sum of the three components are shown with the
  orange solid line. The black dotted line represents the upscattered
  CMB photons without any cooling from stars and dust. The red solid
  line show the CRp induced gamma-ray flux. The black arrow show the
  differential upper limit from Fermi \cite{2010ApJ...717L..71A}. All
  fluxes are calculated for the Fornax cluster within $\rvir$ using a
  point spread function of $0.1\deg$. The boost from Sommerfeld and
  substructures is about 130 and 570, respectively.}
 \label{fig:IR_comp}
\end{minipage}
\end{figure*}

@@

We continue by comparing the estimated differential flux from Fornax
to three other clusters in Fig.~\ref{fig:clu_comp}; the close by well
studied Virgo cluster, the X-ray bright Perseus clusters, and the
massive merging Coma cluster. We find high DM induced gamma-ray fluxes
from both the Fornax and Virgo clusters, which highlight them as
promising targets for indirect DM searches. Note, however, that we
have used spatially unresolved upper limits, which is quite optimistic
for the angular extended Virgo cluster. It is quite striking how
constraining the upper limits for the Fornax cluster is compared to
the other clusters, especially in the $1-10$~GeV energy regime where
Fermi-LAT is most sensitive due to the maximized effective area. The
upper limits for Virgo and Perseus are background contaminated by AGN
activity from M87 and NGC 1267, respectively, and do not gain much
from the increased sensitivity. One should also note that the high CRp
induced flux from both Coma and Perseus is of the same magnitude as
the DM flux from LP models, making it difficult for indirect DM
searches in those clusters. Conversely is the Fornax cluster a great
target for indirect DM studies because the relative low gamma-ray flux
from CRps, absents of an active AGN, and high DM gamma-ray
flux. Actually, the Fermi-LAT 3 year data will be able to constrain
the smallest mass of substructures using the DM BM models. We have
also learned that the closest clusters are superior for DM searches
compared to further away clusters, even though the close by clusters
have a mass that is smaller. In fact, we can understand this by
looking at the luminosity-to-mass scaling relations. The gamma-ray
luminosity from the smooth density profile is given by
\begin{equation}
L_{\gamma,\rmn{sm}} \propto \int \dd V \rho(r)^2 \propto \frac{c^3}
{\left[\log\left(1+c\right)-c/(1+c)\right]} \propto \mvir^{0.83}\,,
\end{equation}
and the total DM luminosity including boost factors for the LP and
the BM models are given by
\begin{eqnarray}
\label{eq:LP_scaling}
L_{\gamma,\rmn{LP}} &=& L_{\gamma,\rmn{smLP}} \B_\rmn{LP} \propto \mvir^{0.73}\qquad
\,\,\,\,\,\,\rmn{and}\\
\label{eq:BM_scaling}
L_{\gamma,\rmn{BM}} &=& L_{\gamma,\rmn{smBM}} \B_\rmn{BM} \propto  \mvir^{1.06}\,,\qquad
\rmn{where} \\
B_\rmn{LP}&=&B_\sfe B_\sub \propto \mvir^{-1/3}\mvir^{0.23} \propto
 \mvir^{-0.1}\,,\\
B_\rmn{BM}&=&B_\sub \propto \mvir^{0.23}\,.
\end{eqnarray}
Hence increasing the mass with a factor two doubles the luminosity,
but increasing the distance by a factor two decrease the flux with a
factor four. It is also usefull to relate the luminosities to a more
model independent variable as the maximum circular velocity given by
\begin{eqnarray}
V_\rmn{max} &=& V_\rmn{circ}(r_\rmn{max})\qquad\,\,\,\,\,\,\,\,\,\rmn{where}\nonumber\\
V_\rmn{circ}(r)&=& \sqrt{\frac{G_\rmn{N}\,M(<r)}{r}}\qquad\rmn{and}\nonumber\\
\frac{\dd V_\rmn{circ}(r)}{\dd r} &=& 0\quad \Longrightarrow\quad r_\rmn{max}\,.
\end{eqnarray}
We find that $V_\rmn{max}=650\,\rmn{km/s}$ of a large $10^{15}\,\msun$
cluster, and $V_\rmn{max}=420\,\rmn{km/s}$ of a small
$4\times10^{13}\,\msun$ group.

@@

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
 \includegraphics[width=0.49\columnwidth]{figures/flux.cluster.Fornax.v9.0.1deg.1.6T.SubMass.SF300.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.cluster.Coma.v9.0.1deg.1.6T.SubMass.SF300.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.cluster.Virgo.v9.0.1deg.1.6T.SubMass.SF300.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.cluster.Perseus.v9.0.1deg.1.6T.SubMass.SF300.IR2.noMW.woGal.eps}
\caption{\it Comparing the emission from different clusters. We show
  the differential flux from four clusters; Fornax (upper left), Coma
  (upper right), Virgo (lower left), and Perseus (lower right). The
  different flux models; CRp induced emission (red solid), leptophilic
  models (LP) include both final state radiation and IC upscattered
  CMB, dust and starlight (green dash-dotted), and benchmark $\Kp$
  models (BM) include continuum emission, and IC upscattered CMB, dust
  and starlight (blue dashed). The arrows have colors matching each
  cluster and show the differential upper limits set by Fermi-LAT in
  the energy ranges $0.2-1$~GeV, $1-10$~GeV, and $10-100$~GeV from
  left to right. For visualization we offset the arrows by increasing
  the mean energy by 30\% in the opposite order as they appear in the
  figure, i.e. starting with Perseus. The flux from the clusters is
  integrated out to $\rvir$ using a point spread function of $0.1\deg$
  and includes the boost from substructures. We find that the lower
  GeV energy regime is most constraining, and induce upper limits on
  boost factors.}
 \label{fig:clu_comp}
\end{minipage}
\end{figure*}

@@

\begin{figure}%[t]
 \includegraphics[width=0.99\columnwidth]{figures/LP.const.diff.v9.0.1deg.1.6T.SubMass.SF300.IR2.noMW.woGal.eps}
\caption{\it Constraining boost factors using flux upper limits. The
  leptophilic (LP) gamma-ray emission is derived within $\rvir$ using
  a point spread function of $0.1\deg$. In order not to overproduce
  the Fermi-LAT differential flux upper limits in the energy interval
  $1-10$~GeV the boost from substructures and Sommerfeld enhancement
  (SFE) is constrained for four clusters; Fornax (red), Virgo
  (orange), M49 (green), and NGC4636 (Blue). The SFE has been rescaled
  from 300 (refer to equation) that is required to explain the
  electron and positron excess observed at Earth with LP DM (black
  arrow). If the DM interpretation is correct, we can constrain the
  smallest size of halos to be larger than $10\,\msun$. However, if
  the smallest size of DM halos is $10^{-6}\,\msun$, we can constrain
  the SFE to about 4 in Fornax, corresponding to a maximum SFE in the
  MW of about 10.}
 \label{fig:boost_const}
\end{figure}

@@

\begin{figure}
 \includegraphics[width=0.99\columnwidth]{figures/SB.v9.1GeV.SF300.SubMass.elmu.eps}
\caption{\it Comparing the surface brightness from different
  clusters. We show the CRp induced emission (red), leptophilic
  emission (green), and emission from the $\Kp$ benchmark model
  (blue). The surface brightness above 1~GeV is derived using a very
  small point spread function ($\psf=10^{-5}\deg$). The
  different linestyles each represent a cluster; Fornax (solid), Coma
  (dotted), Virgo (dashed), and Perseus (dash-dotted). The boost from
  Sommerfeld and substructures is about 130 and 570,
  respectively. Notice the very similar shape and normalization of the DM
  models, while the CRps have a much larger scatter in the profiles.}
 \label{fig:SB_clu}
\end{figure}

@@

\begin{figure}%[t]
 \includegraphics[width=0.99\columnwidth]{figures/SB.resolved.v9.1GeV.SF300.noSuB.vs.SubMass.elmu.eps}
\caption{\it The impact of substructures on surface brightness. We
  show the surface brightness at 1~GeV for the Fornax cluster using a
  very small point spread function of $10^{-5}\deg$ that reveals the
  details of the spatial profiles. The emission induced by CRps is
  denoted by the red solid line, the leptophilic models by green
  lines, and the DM $\Kp$ benchmark model by blue lines. The dashed
  line show the brightness where the boost from substructures is
  included while the substructures are excluded in the dotted
  line. The boost from Sommerfeld and substructures is about 130 and
  570, respectively. Notice the nearly flat profiles when
  substructures are included, and the relative large boost at a
  percent of $\rvir$.}
 \label{fig:SB_sub}
\end{figure}

@@

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
  \includegraphics[width=0.49\columnwidth]{figures/SB.v9.1GeV.SF300.noSuB.elmu.eps}
  \includegraphics[width=0.49\columnwidth]{figures/SB.fornax.v9.1GeV.SF300.noSuB.elmu.eps}
\caption{\it The surface brightness without substructures. We show the
  CRp induced emission (red), leptophilic emission (green), and
  emission from the $\Kp$ benchmark model (blue). The surface
  brightness above 1~GeV is derived using a very small point spread
  function ($\psf=10^{-5}\deg$). LEFT PANEL: compare the
  surface brightness for different clusters; Fornax (solid), Coma
  (dotted), Virgo (dashed), and Perseus (dash-dotted). RIGHT PANEL:
  compare different emission components in Fornax; dotted lines show
  the inverse Compton (IC) upscattered CMB photons, dashed lines show
  the IC upscattered photons from stars and dust, and blue solid line
  show the continuum emission from the $\Kp$ benchmark
  model.}
 \label{fig:SB_clu_nosub}
\end{minipage}
\end{figure*}

@@

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
  \includegraphics[width=0.49\columnwidth]{figures/SB.Fornax.v9.SF300.SubMass.elmu.eps}
  \includegraphics[width=0.49\columnwidth]{figures/SB.Virgo.v9.SF300.SubMass.elmu.eps}
\caption{\it Surface brightness predicted for Cherenkov telescopes at
  high energies. We show the emission above $100$~GeV and include the
  boost from substructures. We use a point spread function of
  $\psf=0.1\deg$ that is typical for both Cherenkov
  telescopes as well as the Fermi-LAT at this energy. Left panel show
  the Fornax cluster and right panel the Virgo cluster. The gamma-ray
  emission is derived for the following components; CRps (red solid),
  continuum emission from the DM $\Kp$ benchmark model (blue
  dash-dotted), as well as final state radiation (green dashed) and
  inverse Compton upscattered dust and starlight (green dotted) from
  leptophilic DM.}
 \label{fig:SB_IACTs}
\end{minipage}
\end{figure*}

\section{Predictions and observational limits}

In this section, we compute the expected $\gamma$-ray fluxes from DM
annihilation and CR interactions of the brightest clusters of the X-ray flux
complete sample in the local universe. We confront these predictions to upper
limits obtained by Fermi 1.5-year data and conclude on the viability of the
underlying models and perspectives for the next years of Fermi observations.

\begin{figure*}
  \includegraphics[width=0.99\columnwidth]{figures/MLscaling.100M.eps}
  \includegraphics[width=0.99\columnwidth]{figures/MLscaling.10G.eps}
  \caption{\it Scaling relations of the cluster's virial mass $M_{200}$ and the
    $\gamma$-ray luminosity for energies $E_\gamma>100$~MeV (left) and
    $E_\gamma>10$~GeV (right).  Shown are the relations for the CR-induced
    emission that is dominated by pion decay resulting from hadronic CR
    interactions (solid), the leptophilic model of DM annihilations that
    includes the scaling of the substructure and Sommerfeld boost factors with
    custer mass from Equation~(\ref{eq:LP_scaling}) (dash-dotted) as well as the
    benchmark model $K'$ of DM annihilation that includes the scaling of the
    substructure boost with cluster mass of Equation~(\ref{eq:BM_scaling})
    (dashed). Note that the scaling of this particular benchmark model is
    indicative for all (non-Sommerfeld enhanced) DM annihilation models while
    the normalization depends on the particular cross-section and neyralino
    masses.}
\label{fig:lum_mass_scaling}
\end{figure*}

First we focus on $\gamma$-ray flux-cluster mass scaling relations for DM
annihilation and CR induced emission in Fig.~\ref{fig:lum_mass_scaling}.  The
mass scaling of the substructure boost steepens the intrinsically shallower DM
annihilation relation of benchmark models (Equation~(\ref{eq:BM_scaling})) while
this steepening is in fact over-compensated by the inverse mass scaling of the
Sommerfeld boost in leptophilic models that prefers smaller halos with a larger
ratio of $c/v_\mathrm{circ}$ (Equation~(\ref{eq:LP_scaling})). In contrast, the
CR scaling relation is considerably steeper as shown by reference
\cite{2010MNRAS.409..449P}. [ANDERS: add CR scaling formula] This implies
already a general strategy to minimize the CR-induced foreground for DM
annihilation and argues for very nearby groups. We note that these simulations
do not include AGn feedback that is thought to furthermore reduce the baryon
fraction in groups relative to that in clusters \cite{2008ApJ...687L..53P}. The
resulting smaller target density for hadronic CR interactions will furthermore
steepen the $L_\gamma-M$ relation of CR induced emission, making the case for
groups even stronger.

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
  \includegraphics[width=0.99\columnwidth]{figures/Fermi.comp.DM.eps}
  \caption{\it Fermi flux upper limits are contrasted to the predicted DM
    gamma-ray flux. We show the mean differential flux in the energy range
    $E_\gamma=1-10$~GeV for 32 clusters. The Fermi-LAT upper limits are shown
    with black arrows. The predicted fluxes are derived from the dominant
    inverse Compton-scattering of CMB photons in a leptophilic DM model (red),
    and the continuum emission from the DM $\Kp$ benchmark model
    (blue). Assuming a boost factor due to substructures that has a mass
    spectrum extending down to Earth masses, the expected leptophilic fluxes are
    ruled out by lower upper limits in 14 of the clusters, with the strongest
    constraints set by Fornax and M49. We can not yet constrain the benchmark
    models, although our predictions are approaching the upper limits and will
    allow to test these in the upcoming years with better data.}
 \label{fig14}
\end{minipage}
\end{figure*}

Figure~\ref{fig14} compares Fermi upper limits on the $\gamma$-ray flux with
predictions of DM annihilation fluxes in the leptophilic and benchmark models.
To obtain the $\gamma$-ray fluxes, we scale the virial mass of clusters in the
extended HIFLUGCS catalogue \cite{2002ApJ...567..716R} with our scaling
relations of Equations~(\ref{eq:BM_scaling}) and (\ref{eq:LP_scaling}).  We
assume a boost factor due to substructures that has a constant contribution per
decade in substructure mass and has a mass spectrum extending down to Earth
masses. Fermi limits rule out the leptophilic models in their present form with
the mentioned assumptions in 14 clusters, and limit the boost from Sommerfeld
enhancement to less than 4 [?] in Fornax. Assuming universality, this limits the
Sommerfeld boost in the MW to less than 10 (see Fig.~\ref{fig:boost_const}).
The flux level of the Fermi limits are an indirect measure of the ambient
background flux in the $\gamma$-ray sky and/or the presence of strong point
sources such as AGN in the immediate vicinity of the cluster position. Hence, the ratio of
the Fermi upper limits to the predicted annihilation fluxes,
$F_{\mathrm{UL}}/F_{\mathrm{scal}}$, is a good indication of the best cluster
candidates for indirect DM experiments. We identify Fornax, M49, NGC4636, and
Virgo to be prime candidates. Note however, that Virgo extends of $7\deg$ over
the sky which implies a lower surface brightness and lower signal-to-noise.
Fermi limits on individual clusters are expected to improve as $\sqrt{T/1.5
  \mathrm{yr}}$, where $T$ is the total elapsed time of the Fermi mission. We
emphasize that the very inhomogeneous distribution of $F_{\mathrm{UL}}/
F_{\mathrm{scal}}$ makes it unlikely to dramatically improve the limits through
a stacking analysis as the weak signal will dilute the average constraints.
Assuming the projected sensitivity of the future Cherenkov telescope array (CTA)
of [ANDERS: value?], we note that it will be difficult to detect the DM
annihilation signal from clusters without a Sommerfeld boost by Cherenkov
telescopes. This is because the boost from substructures is extended while the
sensitivity of Cherenkov telescopes drops linearly with radius outside the point
spread function.

>>>>

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
 \includegraphics[width=0.49\columnwidth]{figures/flux.BMcompI.v10.0.1deg.1.6T.SubMass.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.BMcompJ.v10.0.1deg.1.6T.SubMass.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.BMcompK.v10.0.1deg.1.6T.SubMass.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.BMcompJs.v10.0.1deg.1.6T.SubMass.IR2.noMW.woGal.eps}
\caption{\it Comparing the differential flux from different models: we
  show the continuum emission from DM benchmark (BM) models (blue
  dashed), electrons and positrons from DM BM models that inverse
  Compton upscattered both CMB photons (orange dash-dotted) as well as
  dust and star photons (green dotted), and CRp induced gamma-ray
  emission (red solid). Each panel is associated with an individual DM
  BM model; upper left $\Ip$, lower left $\Kp$, upper right $\Jp$, and
  lower right $\Js$. The emission is calculated for the Fornax cluster
  using a point spread function of $0.1\deg$ and a boost from
  substructures of 570. We find bright prospects for detecting the
  continuum emission that is dominating the total emission in the GeV
  energy range. Also note that, above 100~MeV the total inverse
  Compton emission is at least a factor 10 smaller than the emission
  expected from both the continuum emission and the emission coming
  from CRps.}
 \label{fig:diff_BM}
\end{minipage}
\end{figure*}

@@

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
\includegraphics[width=0.49\columnwidth]{figures/flux.IRcomp.v10.0.1deg.1.6T.SubMass.elmu.SF300.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.IRcomp.BMv10.0.1deg.SubMass.noMW.woGal.eps}
\caption{\it Comparing the flux from different inverse Compton
  upscattered radiation fields. We show the differential inverse
  Compton emission induced by leptophilic DM in the left panel and by
  the $\Kp$ benchmark model in the right panel. The contribution from
  each individual radiation field from top line to bottom line; CMB
  (green dashed-double-dotted), dust (blue dashed), stars (purple
  dashed-dotted). The sum of the three components are shown with the
  orange solid line. The black dotted line represents the upscattered
  CMB photons without any cooling from stars and dust. The red solid
  line show the CRp induced gamma-ray flux. The black arrow show the
  differential upper limit from Fermi \cite{2010ApJ...717L..71A}. All
  fluxes are calculated for the Fornax cluster within $\rvir$ using a
  point spread function of $0.1\deg$. The boost from Sommerfeld and
  substructures is about 130 and 570, respectively.}
 \label{fig:IR_comp}
\end{minipage}
\end{figure*}

@@

We continue by comparing the estimated differential flux from Fornax
to three other clusters in Fig.~\ref{fig:clu_comp}; the close by well
studied Virgo cluster, the X-ray bright Perseus clusters, and the
massive merging Coma cluster. We find high DM induced gamma-ray fluxes
from both the Fornax and Virgo clusters, which highlight them as
promising targets for indirect DM searches. Note, however, that we
have used spatially unresolved upper limits, which is quite optimistic
for the angular extended Virgo cluster. It is quite striking how
constraining the upper limits for the Fornax cluster is compared to
the other clusters, especially in the $1-10$~GeV energy regime where
Fermi-LAT is most sensitive due to the maximized effective area. The
upper limits for Virgo and Perseus are background contaminated by AGN
activity from M87 and NGC 1267, respectively, and do not gain much
from the increased sensitivity. One should also note that the high CRp
induced flux from both Coma and Perseus is of the same magnitude as
the DM flux from LP models, making it difficult for indirect DM
searches in those clusters. Conversely is the Fornax cluster a great
target for indirect DM studies because the relative low gamma-ray flux
from CRps, absents of an active AGN, and high DM gamma-ray
flux. Actually, the Fermi-LAT 3 year data will be able to constrain
the smallest mass of substructures using the DM BM models. We have
also learned that the closest clusters are superior for DM searches
compared to further away clusters, even though the close by clusters
have a mass that is smaller. In fact, we can understand this by
looking at the luminosity-to-mass scaling relations. The gamma-ray
luminosity from the smooth density profile is given by
\begin{equation}
L_{\gamma,\rmn{sm}} \propto \int \dd V \rho(r)^2 \propto \frac{c^3}
{\left[\log\left(1+c\right)-c/(1+c)\right]} \propto \mvir^{0.83}\,,
\end{equation}
and the total DM luminosity including boost factors for the LP and
the BM models are given by
\begin{eqnarray}
L_{\gamma,\rmn{LP}} &=& L_{\gamma,\rmn{smLP}} \B_\rmn{LP} \propto \mvir^{0.73}\qquad
\,\,\,\,\,\,\rmn{and}\\
L_{\gamma,\rmn{BM}} &=& L_{\gamma,\rmn{smBM}} \B_\rmn{BM} \propto  \mvir^{1.06}\,,\qquad
\rmn{where} \\
B_\rmn{LP}&=&B_\sfe B_\sub \propto \mvir^{-1/3}\mvir^{0.23} \propto
 \mvir^{-0.1}\,,\\
B_\rmn{BM}&=&B_\sub \propto \mvir^{0.23}\,.
\end{eqnarray}
Hence increasing the mass with a factor two doubles the luminosity,
but increasing the distance by a factor two decrease the flux with a
factor four. It is also usefull to relate the luminosities to a more
model independent variable as the maximum circular velocity given by
\begin{eqnarray}
V_\rmn{max} &=& V_\rmn{circ}(r_\rmn{max})\qquad\,\,\,\,\,\,\,\,\,\rmn{where}\nonumber\\
V_\rmn{circ}(r)&=& \sqrt{\frac{G_\rmn{N}\,M(<r)}{r}}\qquad\rmn{and}\nonumber\\
\frac{\dd V_\rmn{circ}(r)}{\dd r} &=& 0\quad \Longrightarrow\quad r_\rmn{max}\,.
\end{eqnarray}
We find that $V_\rmn{max}=650\,\rmn{km/s}$ of a large $10^{15}\,\msun$
cluster, and $V_\rmn{max}=420\,\rmn{km/s}$ of a small
$4\times10^{13}\,\msun$ group.

@@

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
 \includegraphics[width=0.49\columnwidth]{figures/flux.cluster.Fornax.v10.0.1deg.1.6T.SubMass.SF300.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.cluster.Coma.v10.0.1deg.1.6T.SubMass.SF300.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.cluster.Virgo.v10.0.1deg.1.6T.SubMass.SF300.IR2.noMW.woGal.eps}
\includegraphics[width=0.49\columnwidth]{figures/flux.cluster.Perseus.v10.0.1deg.1.6T.SubMass.SF300.IR2.noMW.woGal.eps}
\caption{\it Comparing the emission from different clusters. We show
  the differential flux from four clusters; Fornax (upper left), Coma
  (upper right), Virgo (lower left), and Perseus (lower right). The
  different flux models; CRp induced emission (red solid), leptophilic
  models (LP) include both final state radiation and IC upscattered
  CMB, dust and starlight (green dash-dotted), and benchmark $\Kp$
  models (BM) include continuum emission, and IC upscattered CMB, dust
  and starlight (blue dashed). The arrows have colors matching each
  cluster and show the differential upper limits set by Fermi-LAT in
  the energy ranges $0.2-1$~GeV, $1-10$~GeV, and $10-100$~GeV from
  left to right. For visualization we offset the arrows by increasing
  the mean energy by 30\% in the opposite order as they appear in the
  figure, i.e. starting with Perseus. The flux from the clusters is
  integrated out to $\rvir$ using a point spread function of $0.1\deg$
  and includes the boost from substructures. We find that the lower
  GeV energy regime is most constraining, and induce upper limits on
  boost factors.}
 \label{fig:clu_comp}
\end{minipage}
\end{figure*}

@@

\begin{figure}%[t]
 \includegraphics[width=0.99\columnwidth]{figures/LP.const.diff.v10.0.1deg.1.6T.SubMass.SF300.IR2.noMW.woGal.eps}
\caption{\it Constraining boost factors using flux upper limits. The
  leptophilic (LP) gamma-ray emission is derived within $\rvir$ using
  a point spread function of $0.1\deg$. In order not to overproduce
  the Fermi-LAT differential flux upper limits in the energy interval
  $1-10$~GeV the boost from substructures and Sommerfeld enhancement
  (SFE) is constrained for four clusters; Fornax (red), Virgo
  (orange), M49 (green), and NGC4636 (Blue). The SFE has been rescaled
  from 300 (refer to equation) that is required to explain the
  electron and positron excess observed at Earth with LP DM (black
  arrow). If the DM interpretation is correct, we can constrain the
  smallest size of halos to be larger than $10\,\msun$. However, if
  the smallest size of DM halos is $10^{-6}\,\msun$, we can constrain
  the SFE to about 4 in Fornax, corresponding to a maximum SFE in the
  MW of about 10.}
 \label{fig:boost_const}
\end{figure}

@@

\begin{figure}
 \includegraphics[width=0.99\columnwidth]{figures/SB.v10.1GeV.SF300.SubMass.elmu.eps}
\caption{\it Comparing the surface brightness from different
  clusters. We show the CRp induced emission (red), leptophilic
  emission (green), and emission from the $\Kp$ benchmark model
  (blue). The surface brightness above 1~GeV is derived using a very
  small point spread function ($\psf=10^{-5}\deg$). The
  different linestyles each represent a cluster; Fornax (solid), Coma
  (dotted), Virgo (dashed), and Perseus (dash-dotted). The boost from
  Sommerfeld and substructures is about 130 and 570,
  respectively. Notice the very similar shape and normalization of the DM
  models, while the CRps have a much larger scatter in the profiles.}
 \label{fig:SB_clu}
\end{figure}

@@

\begin{figure}%[t]
 \includegraphics[width=0.99\columnwidth]{figures/SB.resolved.v10.1GeV.SF300.noSuB.vs.SubMass.elmu.eps}
\caption{\it The impact of substructures on surface brightness. We
  show the surface brightness at 1~GeV for the Fornax cluster using a
  very small point spread function of $10^{-5}\deg$ that reveals the
  details of the spatial profiles. The emission induced by CRps is
  denoted by the red solid line, the leptophilic models by green
  lines, and the DM $\Kp$ benchmark model by blue lines. The dashed
  line show the brightness where the boost from substructures is
  included while the substructures are excluded in the dotted
  line. The boost from Sommerfeld and substructures is about 130 and
  570, respectively. Notice the nearly flat profiles when
  substructures are included, and the relative large boost at a
  percent of $\rvir$.}
 \label{fig:SB_sub}
\end{figure}

@@

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
  \includegraphics[width=0.49\columnwidth]{figures/SB.v10.1GeV.SF300.noSuB.elmu.eps}
  \includegraphics[width=0.49\columnwidth]{figures/SB.fornax.v10.1GeV.SF300.noSuB.elmu.eps}
\caption{\it The surface brightness without substructures. We show the
  CRp induced emission (red), leptophilic emission (green), and
  emission from the $\Kp$ benchmark model (blue). The surface
  brightness above 1~GeV is derived using a very small point spread
  function ($\psf=10^{-5}\deg$). LEFT PANEL: compare the
  surface brightness for different clusters; Fornax (solid), Coma
  (dotted), Virgo (dashed), and Perseus (dash-dotted). RIGHT PANEL:
  compare different emission components in Fornax; dotted lines show
  the inverse Compton (IC) upscattered CMB photons, dashed lines show
  the IC upscattered photons from stars and dust, and blue solid line
  show the continuum emission from the $\Kp$ benchmark
  model.}
 \label{fig:SB_clu_nosub}
\end{minipage}
\end{figure*}

@@

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
  \includegraphics[width=0.49\columnwidth]{figures/SB.Fornax.v10.SF300.SubMass.elmu.eps}
  \includegraphics[width=0.49\columnwidth]{figures/SB.Virgo.v10.SF300.SubMass.elmu.eps}
\caption{\it Surface brightness predicted for Cherenkov telescopes at
  high energies. We show the emission above $100$~GeV and include the
  boost from substructures. We use a point spread function of
  $\psf=0.1\deg$ that is typical for both Cherenkov
  telescopes as well as the Fermi-LAT at this energy. Left panel show
  the Fornax cluster and right panel the Virgo cluster. The gamma-ray
  emission is derived for the following components; CRps (red solid),
  continuum emission from the DM $\Kp$ benchmark model (blue
  dash-dotted), as well as final state radiation (green dashed) and
  inverse Compton upscattered dust and starlight (green dotted) from
  leptophilic DM.}
 \label{fig:SB_IACTs}
\end{minipage}
\end{figure*}

\section{Predictions and observational limits}
Model predictions

\begin{figure}
  \includegraphics[width=0.99\columnwidth]{figures/MLscaling.eps}
  \caption{\it Luminosity to mass scaling relations. We show the
    luminosity from; CRp induced emission (red solid), a leptophilic
    model that IC upscatter CMB photons (green dash-dotted), mass
    dependence of SD cooling have been neglected, actually a factor
    few difference for groups. factor 2 for Fornax, and small
    difference for large clusters}
 \label{fig:lum_mass_scaling}
\end{figure}

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
  \includegraphics[width=0.99\columnwidth]{figures/Fermi.comp.DM.eps}
  \caption{\it Fermi flux upper limits contrasted to the predicted DM
    gamma-ray flux. We show the mean differential flux in the energy
    range $E_\gamma=1-10$~GeV for 32 clusters. The Fermi-LAT upper
    limits are shown with black arrows. The predicted fluxes are
    derived from both a leptophilic DM model that result in inverse
    Compton upscattered CMB photons (red), and the continuum emission
    from the DM $\Kp$ benchmark model (blue). The leptophilic fluxes
    are ruled out by 14 of the clusters, with the strongest
    constraints set by Fornax and M49. We can not constrain the
    benchmark models, although our predictions are approaching the
    upper limits.}
 \label{fig14}
\end{minipage}
\end{figure*}
 difficult to detect DM without a Sommerfeld boost with Cherenkov
 telescopes since the boost from substructures is extended while the
 sensitivity of Cherenkov telescopes drops linearly with radius
 outside the point spread function (CHECK)

we rule out the leptophilic models in present form in 14 clusters,
 and limit the boost from Sommerfeld enhancement in the MW to less
 than 10

mention prime cluster candidates, LP under pressure, Fornax start
being constraining, since only few clusters close to UL stacking will
not increase the DM limits much

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
  \includegraphics[width=0.99\columnwidth]{figures/Fermi.comp.CR.diff.eps}
  \caption{\it Fermi-LAT flux upper limits contrasted to the
    predictions by the semi-analytic Pinzke and Pfrommer model for CRp
    induced gamma-ray emission. We show the mean differential flux in
    the energy range $E_\gamma=1-10$~GeV for 32 clusters. The Fermi
    upper limits are shown with black arrows. The blue boxes show the
    gamma-ray emission from CRp induced $\pi^0$\:s decaying, where the
    upper bound shows the estimates for an optimistic model and the
    lower bound a more conservative model excluding point sources (the
    realistic case is more likely in between, see
    \cite{2010MNRAS.409..449P} for details). Note that there is
    currently no tension between our predictions and upper limits from
    Fermi, although Virgo, Norma, and Coma are close and will in the
    coming years enforce constraints on the hadronic models.}
 \label{fig15}
\end{minipage}
\end{figure*}

the tightest limits on DM and CRps from Fermi are in the 1-10 GeV
 regime, where Fornax, M49, and NGC4636 are the most constraining
 "clusters" for the DM and Virgo, Coma and Norma are very close to the
 predictions for the CRps

using beta profiles, discrepancy with Norma cluster in Ackermann gone,
Integrated flux similar

AGN M87, effective area better in 1 GeV range
